---
title: "7.8.3 GAMs"
author: "Team 13"
date: "2/17/2021"
output:
  pdf_document: default
  html_document: default
---
```{r}
rm(list=ls())
library(ISLR)
library(splines)
attach(Wage)
```
GAM -- Generalized Additive Models

We first fit a GAM to predict wage using natural spline functions of year and age, treating education as a qualitative predictor. 

```{r}
gam1 = lm(wage~ns(year,4)+ns(age,5)+education, data=Wage)
```
Since this is just a big linear regression model using an appropriate choice of basis functions, we can simply do this using the lm() function.


We now fit the model using smoothing splines rather than natural splines. In order to fit more general sorts of GAMS, using smoothing splines or other components that cannot be expressed in terms of basis functions and then fit using least squares regression, we will need to use the gam library in R.

Year should have 4 degrees of freedom, age should have 5 degrees of freedom, education is qualitative -- leave it as is. 

```{r}
library(gam)
gam.m3=gam(wage~s(year,4)+s(age,5)+education, data=Wage)
par(mfrow=c(1,3))
plot(gam.m3, se=TRUE, col="blue")

```

The general plot() function recognizes gam.m3 as an object of class gam, and invokes the appropriate plot.gam()method. 
Even gam1 is not of class gam but rather of class lm, we can still use plot.gam() on it. Plot.Gam reproduces figure 7.11 on page 284, instead of the general plot function.
```{r}
plot.Gam(gam1, se=TRUE, col='red')
```
Note that the proper syntax on the current version R is plot.Gam() instead of lower case plot.gam.


In these plots, the function of year looks rather linear. We can perform a series of ANOVA tests in order to determin which of these three models is best: a GAM that excludes year, a GAM that usese a linear function of year, or a GAM that uses a spline function of year.


```{r}
gam.m1<-gam(wage~s(age,5)+education, data=Wage)
gam.m2<-gam(wage~year+s(age,5)+education, data=Wage)
anova(gam.m1, gam.m2, gam.m3, test="F")

```

From the ANOVA tests above we find that there is compelling evidence that a GAM with a linear function of year is better than a GAM that does not include year at all (p-value=0.00014). However, there is no evidence that a non-linear function of year is needed(p-value=0.349). In other words, based on the results of this ANOVA, m2 is preferred.

```{r}
summary(gam.m3)
```

The p-value of year and age correspond to a null hypothesis of a linear relationship versus the alternative of a non-linear relationship. The large p-value for year reinforces our conclusion from the ANOVA test that a linear function is adequate for this term. However, there is ver clear evidence that a non-linear term is required for age. 

We can make predictions from gam objects, just like from lm objects, using the predict() method for the class gam. Here we make predictions on the training set for the best model -- m2. 
```{r}
preds=predict(gam.m2, newdata=Wage)
```

We can also use local regression fits as building blocks in a GAM, using the lo() function.
```{r}
gam.lo=gam(wage~s(year, df=4)+lo(age,span=0.7)+education, data=Wage)
plot.Gam(gam.lo, se=TRUE, col="green")
```

We can also use the lo() function to create interactions before calling the gam() function.
```{r}
gam.lo.i=gam(wage~lo(year,age,span=0.5)+education, data=Wage)
```
It fits a two-term model, in which the first term is an interaction between year and age, fit by a local regression surface. We can plot the resulting two-dimensional surface if we first install the akima package.

```{r}
library(akima)
plot(gam.lo.i)
```

In order to fit a logistic regression GAM, we once again use the I() function in constructing the binary response variable, and set family=binomial.
```{r}
gam.lr=gam(I(wage>250)~year+s(age,df=5)+education,family=binomial, data=Wage)
par(mfrow=c(1,3))
plot(gam.lr,se=T,col='green')

```
It is easy to see that there are no high earners in the <HS category
```{r}
table(education, I(wage>250))
```
Hence we fit a logistic regression GAM using all but this category. This provides more sensible results.

```{r}
gam.lr.s=gam(I(wage>250)~year+s(age,df=5)+education, family=binomial, data=Wage, subset=(education!="1. < HS Grad"))
plot(gam.lr.s,se=T,col="green")
```
