---
title: "Applied Exercise 11"
author: "Team 13 - Xander Giarracco, Xuanxiong Zhen, Hanzheng Li, Henry Tazewell"
date: "2/16/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ISLR)
attach(Hitters)
```

# Applied Exercise 11

GAMs are generally fit using a backfitting approach. This exercise explores the
idea behind backfitting by approximating multiple linear regression using 
simple linear regression in an iterative process.

Suppose that we would like to perform multiple linear regression, but
we do not have software to do so. Instead, we only have software
to perform simple linear regression. Therefore, we take the following
iterative approach: we repeatedly hold all but one coefficient estimate
fixed at its current value, and update only that coefficient
estimate using a simple linear regression. The process is continued until
convergence—that is, until the coefficient estimates stop changing.


Generate a response Y and two predictors X1 and X2

```{r data}
y<-Hitters$Runs
x1<-Hitters$Hits
x2<-Hitters$AtBat
```

Here is the model we are trying to fit:

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \epsilon$$


Initialize $\beta_1$ to take on a value of your choice, in this example we will
arbitrarily choose 1.
```{r beta}
beta1<-1
```

We then rearrange the model such that $\beta_0$ and $\beta_2$ are held constant:

$$Y-\hat{\beta_1} X_1 = \beta_0 + \beta_2 X_2 + \epsilon$$
 
Keeping $\beta_1$ fixed, fit the model:

```{r beta2}
a<-y-beta1*x1
beta2<-lm(a~x2)$coef[2]
```

Using the same idea as above, we then rearrange the model such that $\beta_0$ and $\beta_1$ are held constant:

$$Y-\hat{\beta_2} X_2 = \beta_0 + \beta_1 X_1 + \epsilon$$

Keeping $\beta_2$ fixed, fit the model:

```{r beta1}
a<-y-beta2*x2
beta1<-lm(a~x1)$coef[2]
```

Write a for loop to repeat the process 100 times. Report the
estimates of $\beta_0$, $\beta_1$, and $\beta_2$ at each iteration of the for loop.
We can view the print out below:

```{r loop}
betas<-data.frame(beta0=numeric(),beta1=numeric(),beta2=numeric())
x<-1:100
for(i in x) {
  a<-y-beta1*x1
  beta2<-lm(a~x2)$coef[2]
  
  a<-y-beta2*x2
  beta1<-lm(a~x1)$coef[2]
  
  beta0<-lm(a~x1)$coef[1]
  
  betas<-rbind(betas, data.frame(beta0=beta0,beta1=beta1,beta2=beta2))
  print(betas[i,])
}
```

Plot the each $\beta$ in a different color. Then create the proper multiple linear
regression model using lm() and use abline() to plot the values of each 
new $\beta$ in the same view.

```{r plots}
plot(x,betas$beta0,type="l",col="red",lty=2,xlab='iteration',ylab='betas')
lines(x,betas$beta1,type="l",col="green",lty=2)
lines(x,betas$beta2,type="l",col="blue",lty=2)

mod<-lm(y~x1+x2)
summary(mod)

abline(mod$coef[1],0, col='pink',lty=3)
abline(mod$coef[2],0,col='lightgreen',lty=3)
abline(mod$coef[3],0,col='lightblue',lty=3)

```


On this data set, it took roughly 70 backfitting iterations to obtain a near perfect
approximation for all three of the multiple regression coefficient estimates.